RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-31 casestudies.tex
yes

We present two example applications of the capabilities provided by RCloud
based on synthetic data and applications.\footnote{Unfortunately, we could
not show production notebooks because of concerns about revealing proprietary
or confidential information.} The main pont of the examples is that they are
representative of a progression through the visual analytics development
and deployment process that we aim to support.

@@

\begin{figure}
\includegraphics[width=\linewidth]{fig/casestudy1/casestudy1.pdf}
\caption{\label{sec:stockvis}Iterations of a stock ticker
  visualization, after an example by Hadley Wickham.  A
  simple, static visualization of the closing price of a single ticker
  is progressively developed into a configurable display suitable
  for dashboarding, then into an interactive visualization of the
  volatility and volume of two stocks, then finally into an API
  for data access by other RCloud notebooks. All notebooks
  in this example can be loaded as web pages. When a notebook
  corresponding to a function call is displayed a web page,
  its associated documentation is displayed.}
\end{figure}

@@

The webpage that produces that visualization includes a link back to
the generating source code, from which a user can fork the notebook
and, in this case, add a configurable ticker based on the URL of
the notebook. Notice how the notebook changes very little.

From there, the notebook author (or another user) may decide to
provide convenient access to the data apart from its visualization.
This is achieved by creating a notebook that defines a function.
This notebook then becomes a \emph{subroutine} for other notebooks,
and is version-controlled like other notebooks.

>>>>

We present two example applications in RCloud based on synthetic data
and applications.\footnote{Unfortunately, we could not show production
notebooks that may contain proprietary or confidential information.}
The main point of the examples is to show key steps in the visual
analytics development and deployment process that we aimed to support.

@@

\begin{figure}
\includegraphics[width=\linewidth]{fig/casestudy1/casestudy1.pdf}
\caption{\label{sec:stockvis}Iterations of a stock ticker
  visualization, based on an example by Hadley Wickham. A
  simple, static visualization of the closing price of a single stock
  is progressively developed into a configurable display suitable
  for dashboarding, then into an interactive visualization for 
  comparing the volatility and volume of two stocks, and finally
  into an API for data access by other RCloud notebooks. The notebooks
  in this example can all be loaded as web pages. When a notebook
  corresponding to a function call is displayed a web page,
  its associated documentation is displayed.}
\end{figure}

@@

The webpage that produces that visualization includes a link back to
the generating source code, from which a user can fork the notebook
and, in this case, add a configurable ticker based on the URL of
the notebook. Notice that the notebook changes very little.

From there, the notebook author or a collaborator may decide to
provide convenient access to the data, apart from its visualization.
This is done by creating a notebook that defines a function.
That notebook then becomes a \emph{subroutine} for other notebooks,
and is version-controlled like other notebooks.

### 2014-03-31 discussion.tex
yes

Experience with deploying RCloud in a community of data analysts
(hackers in the classification of Heer and Agrawala)
has provided some insight about how the proposed requirements
were met and to what extent they were appropriate.
This experience may be informative to researchers and implementers
of future systems supporting collaborative visual analytics.

>>>>

Experience with deploying RCloud in a community of data analysts
(``hackers'') has provided some insight about how the proposed
requirements were met and to what extent they were appropriate.
This experience may be informative to researchers and implementers
of future systems supporting collaborative visual analytics.

### 2014-03-31 introduction.tex
yes

% Why does this not work?
There are two problems with the current practice. First, there are
gaps in the workflow: Exploratory Data Analysis (EDA) is done with one
set of tools, and automated reports and deployments with another. This
makes transferring results between EDA and production (in both
directions) slow, expensive, and error-prone. Coordination is done
through meetings and may take months to implement.

Second, EDA environments often assume a single-developer perspective,
while data scientist teams could greatly benefit from convenient searching
and sharing of scripts, data sets, data feeds, experiments, annotations,
and automated analyses and recommendations. These capabilities have had
tremendous impact on information retrieval in general but are well
beyond what existing visual analytics environments provide, or what
is found in today's repositories and version control
systems. \todo{add cites to reco system work}

In addition, take the recent trend in interactive visualization
technologies. Let us consider prominent designers and developers of
open source software, Jeffrey Heer, and Hadley Wickham. Heer has
written (or helped write) Prefuse, Flare, Protovis, d3, and Vega, in
that order. All but the first of these libraries target web
technologies, and all but the second use open standards. Wickham has
written ggplot, ggplot2, and ggvis. ggplot and
ggplot2 generate static graphics, while ggvis uses Shiny and Vega to
provide easy interactive visualizations in R, by opening a separate
web server for each call. \todo{Add cites for all of these}

@@

We see the present state-of-the-art in environments for data science
teams as tremendously powerful, but ultimately fragmented. Rather than
design a completely new solution that risks even further
fragmentation, we wish to find and remedy the shortcomings of data
science as we see it practiced today. The following requirements
summarize what we consider to be the main pain points:

@@

\item Technology transfer. In our experience, development
of new analyses and visualizations is done by \emph{hackers}
and \emph{scripters} (where here are use Kandel et al.'s
terminology~\cite{Kandel:2012:EDA}) who are, by definition, familiar
to some extent with programming and software development.
On the other hand, \emph{application
users} are the ones extracting value from the data using the tools
developed by hackers and scripters. Traditionally, the communication
between these two world is done by IT staff. In an environment where
business needs change possibly many times a week, manual IT support
does not scale: there needs to exist
infrastructure to connect the two worlds.
%% The two classes of users
%% have quite different perspectives but we
%% infrastructure for continuous deployment of visualizations.

>>>>

% Why does this not work?
There are two problems with the current practice. First, there are
gaps in the workflow: Exploratory Data Analysis (EDA) is done with one
set of tools, and automated reports and deployments with another. This
makes transferring results between EDA and production (in both
directions) slow, expensive, and error-prone. Coordination is done
through meetings and may take months to carry out.

Second, EDA environments often assume a single-developer perspective,
while data scientist teams could greatly benefit from convenient searching
and sharing of scripts, data sets, data feeds, experiments, annotations,
and automated analyses and recommendations. These capabilities have had
tremendous impact on information retrieval in general but are well
beyond what existing visual analytics environments provide, or what
is found in today's repositories and version control
systems.\cite{Koren:2009:FIT}

In addition, consider recent trends in interactive visualization
technologies in the open source community. Projects led by Jeff Heer
include Prefuse, Flare, Protovis, d3, and Vega, in that order.
All but the first of these libraries take advantage of web technologies,
and all but the second rely on open standards. Hadley Wickham is the
author of ggplot, ggplot2, and ggvis. ggplot and ggplot2 generate
static graphics, while ggvis provides convenient interactive visualizations
in R, using the R package Shiny and the Vega abstract visualization grammar,
opening a separate web server for each call.

@@

Today's state-of-the-art in environments for data science teams are extremely
powerful, but also fragmented. Instead of creating a new solution that risks
even further fragmentation, we wish to identify and remedy the shortcomings
in current technology for the practice of data science. The following
requirements summarize the main pain points:

@@

\item Technology transfer. In our experience, development
of new analyses and visualizations is done by \emph{hackers}
and \emph{scripters} (using Kandel et al.'s
terminology~\cite{Kandel:2012:EDA}) who are, by definition, familiar
to some extent with programming and software development.
On the other hand, \emph{application
users} are the ones extracting value from the data using the tools
developed by hackers and scripters. Traditionally, the communication
between these two world is done by IT staff. In an environment where
business needs change possibly many times a week, manual IT support
does not scale: there needs to exist
infrastructure to connect the two worlds.
%% The two classes of users
%% have quite different perspectives but we
%% infrastructure for continuous deployment of visualizations.
